A point-level trend model for the North American Breeding Bird Survey that corrects for detectability

### 2024-08-30 article/article.tex
table

\usepackage{array}

@@

\begin{table}[ht]\label{table:params}
	\centering
	\caption{Parameters in the BBS models.}
		\begin{tabular}[t]{l>{\raggedright}p{0.5\linewidth}>{\raggedright\arraybackslash}p{0.3\linewidth}}
		\hline
		Parameter or Datum&Description&Model\\
		\hline
		$i$&Stratum index&Equation \ref{bbs}, \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$j$&Point/stop index&Equation \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$J$&Route index&Equation \ref{bbs}\\
		$t$&Year index&Equation \ref{bbs}, \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$o$&Observer index&Equation \ref{bbs}, \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$Y_{J,i,t,o}$&Total observed birds at route $J$, stratum $i$, year $t$, by observer $o$&Equation \ref{bbs}\\
		$Y_{j,i,t,o}$&Total observed birds at point $j$, stratum $i$, year $t$, by observer $o$&Equation \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$\alpha_i$&Stratum-level intercept, modelled with an intrinsic conditional autoregressive model \citep{besag_bayesian_1991, smith_spatially_2023}&Equation \ref{bbs}, \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$\delta_{i,t}$&Smoothed temporal component estimating population trajectory through time. Modelled with generalized additive model \citep{smith_north_2021}&Equation \ref{bbs}, \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$\gamma_{i,t}$&Annual year-effect fluctuation term; departure from smoothed GAM \citep{smith_north_2021}&Equation \ref{bbs}, \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$\psi_J$&Route-level intercept&Equation \ref{bbs}\\
		$\psi_j$&Point-level intercept&Equation \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$\omega_o$&Observer-level intercept&Equation \ref{bbs}, \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$\eta$&Smoothed temporal component estimating population trajectory through time. Modelled with generalized additive model \citep{smith_north_2021}&Equation \ref{bbs}, \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		\hline
	\end{tabular}
\end{table}%

@@

$$\log(p) = \log p(\hat{\theta}) + \zeta\kappa^{(p)}$$
We can now see that we have introduced the estimated availability (i.e., $\log p(\hat{\theta})$) and as a result of the Taylor expansion we have introduced the variance propagation term $\zeta\kappa^{(p)}$.
It follows that if we now perform the same expansions to $\log(q)$, we would get that $\log(q) = \log q(\hat{\theta}) + \xi\kappa^{(q)} + O(\hat{\theta} - \theta_0)^2$ with $\xi$ and $\kappa^{(q)}$ similarly defined.

>>>>

\usepackage{array}
\usepackage{longtable}

@@

\begin{table}[ht]\label{table:params}
	\centering
	\caption{Parameters in the BBS models.}
		\begin{tabular}[t]{l>{\raggedright}p{0.5\linewidth}>{\raggedright\arraybackslash}p{0.3\linewidth}}
		\hline
		Parameter or Datum&Description&Model\\
		\hline
		$i$&Stratum index&Equation \ref{bbs}, \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$j$&Point/stop index&Equation \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$J$&Route index&Equation \ref{bbs}\\
		$t$&Year index&Equation \ref{bbs}, \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$o$&Observer index&Equation \ref{bbs}, \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$Y_{J,i,t,o}$&Total observed birds at route $J$, stratum $i$, year $t$, by observer $o$&Equation \ref{bbs}\\
		$Y_{j,i,t,o}$&Total observed birds at point $j$, stratum $i$, year $t$, by observer $o$&Equation \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$\alpha_i$&Stratum-level intercept, modelled with an intrinsic conditional autoregressive model \citep{besag_bayesian_1991, smith_spatially_2023}&Equation \ref{bbs}, \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$\delta_{i,t}$&Smoothed temporal component estimating population trajectory through time. Modelled with generalized additive model \citep{smith_north_2021}&Equation \ref{bbs}, \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$\gamma_{i,t}$&Annual year-effect fluctuation term; departure from smoothed GAM&Equation \ref{bbs}, \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$\psi_J$&Route-level intercept&Equation \ref{bbs}\\
		$\psi_j$&Point-level intercept&Equation \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$\omega_o$&Observer-level intercept&Equation \ref{bbs}, \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$\eta$&First-year observer effect&Equation \ref{bbs}, \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$I(o,J,t)$&1 if observer is surveying route $J$ in year $t$ for the first time, 0 otherwise&Equation \ref{bbs}\\
		$I(o,j,t)$&1 if observer is surveying point $j$ in year $t$ for the first time, 0 otherwise&Equation \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$p_{j,i,t}$&Probability of availability at point $j$ in stratum $i$ in year $t$&Equation \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$q_{j,i,t}$&Probability of perceptibility at point $j$ in stratum $i$ in year $t$&Equation \ref{bbs-point}, \ref{bbs-detect}, \ref{bbs_varprop}\\
		$\kappa_{j,i,t}^{(p)}$&Design matrix for availability variance propagation; first derivatives of probability of availability \citep{bravington_variance_2021}&Equation \ref{bbs_varprop}\\
		$\kappa_{j,i,t}^{(q)}$&Design matrix for perceptibility variance propagation; first derivatives of probability of perceptibility \citep{bravington_variance_2021}&Equation \ref{bbs_varprop}\\
		$\zeta$&Effects vector for availability variance propagation \citep{bravington_variance_2021}&Equation \ref{bbs_varprop}\\
		$\xi$&Effects vector for perceptibility variance propagation \citep{bravington_variance_2021}&Equation \ref{bbs_varprop}\\
		\hline
	\end{tabular}
\end{table}%

@@

$$\log(p) = \log p(\hat{\theta}) + \zeta\kappa^{(p)}$$
We can now see that we have introduced the estimated availability (i.e., $\log p(\hat{\theta})$) and as a result of the Taylor expansion we have introduced the variance propagation term $\zeta\kappa^{(p)}$.
It follows that if we now perform the same expansions to $\log(q)$, we would get that $\log(q) = \log q(\hat{\theta}) + \xi\kappa^{(q)}$ with $\xi$ and $\kappa^{(q)}$ similarly defined.
